\vspace{-.5cm} Lights: a generalized joint model for high-dimensional multivariate longitudinal data and censored durations \vspace{.5cm}

### 2021-03-21 paper.tex
define p_{\varrho}(y) outside the algo

Now, for the update of $\beta_k^{(w)}$ for $k=0, \ldots, K-1$, we similarly denote
\begin{align*}
  R^{(w)}_{n,k}(\beta_k) = -n^{-1} \sum_{i=1}^n \hat \pi_{ik}^{\theta^{(w)}} \Big[ &\delta_i \hat \E_{\theta^{(w)}}[ \tilde g_{t_i, \gamma_k^{(w)}, \beta_k}^2(b_i) | t_i, \delta_i, y_i] \\
  & - \sum_{j=1}^J \lambda_0^{(w)}(t_j) \hat \E_{\theta^{(w)}}[ \tilde g_{i, t_j, {\gamma_k^0}^{(w)}, \gamma_k^{(w)}, \beta_k}^1(b_i) | t_i, \delta_i, y_i] \ind{\{t_j\leq t_i\}} \\
  & + (y_i \odot \Phi_i^{(w)})^\top \hat \E_{\theta^{(w)}}[ M_{ik} | t_i, \delta_i, y_i] \\
  & - \hat \E_{\theta^{(w)}}[ c_{\phi^{(w)}}(M_{ik}) | t_i, \delta_i, y_i] \Big] 
\end{align*}
based on the quantities involved in $\cQ_n(\theta, \theta^{(w)})$ depending on $\beta_k$, with $\tilde g_{i, t_j, {\gamma_k^0}^{(w)}, \gamma_k^{(w)}, \beta_k}^1$ and $\tilde g_{t_i, \gamma_k^{(w)}, \beta_k}^2$ defined in~\eqref{eq:g-tilde-1-2}. The integration over the survival process has been replaced with a finite sum over the process evaluated at the distinct event times $(t_1, \ldots, t_J)$ with $J \in~\N_+$, since $\lambda_0^{(w)}$ defined in~\eqref{eq:update_lambda_0} is always zero except at observed event times.
The update for $\beta_k^{(w)}$ therefore requires to solve the following convex minimization problem
\begin{equation}
  \label{eq:minimization-beta_k}
  \beta_k^{(w+1)} \in \argmin_{\beta_k \in \R^q} R^{(w)}_{n,k}(\beta_k) + \zeta_{2,k} \norm{\beta_k}_{\text{sg} l_1, \tilde{\eta}}.
\end{equation}
The previous trick we used to handle the non-differentiability of the $\ell_1$ part in the elastic net is no longer sufficient for~\eqref{eq:minimization-beta_k}, since the $\ell_2$ norm in the sparse group lasso is not squared, leading to a non-differentiability at 0 (which is precisely the criterion exploited by the penalty to set some groups of coefficients to exactly 0~\citep{simon2013sparse}). Then, we choose to solve problem~\eqref{eq:minimization-beta_k} using the proximal gradient descent~\citep{boyd2004convex} described in Algorithm~\ref{alg:ISTA}.

@@

\begin{algorithm}[H]
\DontPrintSemicolon
\caption{Proximal gradient descent ISTA$_{\text{sg} l_1}$($v$, $f$, $\tilde \eta$, $\zeta$)}
\label{alg:ISTA}
\KwInput{A vector $v$, a function $f$ (its gradient $\nabla f$), hyper-parameters $\tilde \eta$, $tol$ and $\zeta$}
\KwOutput{$\flat^{\breve{w}^{max}}$}
Define $p_{\varrho}(y) = \argmin_u \Big\{\dfrac{1}{2\varrho} \left\lVert u - \big(y - \varrho  \nabla f(y) \big)\right\rVert_2^2 + \zeta \norm{u}_{\text{sg} l_1, \tilde{\eta}} \Big\}$\\
Initialize $\flat^{(0)}=v$, $\varrho^0 > 0$, some $\epsilon > 1$ \\
\For{$\breve{w} = 0,\ldots,$ until convergence}    
{Backtracking line search : find the smallest nonnegative integer $i^{\breve{w}}$ such that
$f\big(p_{\tilde{\varrho}_{\breve{w}}}(\flat^{(\breve{w})})\big) \leq f(\flat^{(\breve{w})}) + \big<\big(p_{\tilde{\varrho}_{\breve{w}}}(\flat^{(\breve{w})}) - \flat^{(\breve{w})}\big), \nabla f(\flat^{(\breve{w})})\big> + \dfrac{1}{2\tilde{\varrho}_{\breve{w}}}\norm{p_{\tilde{\varrho}_{\breve{w}}}(\flat^{(\breve{w})}) - \flat^{(\breve{w})}}_2^2$ 
with $\tilde{\varrho}_{\breve{w}} = \epsilon^{i^{\breve{w}}}\varrho^{\breve{w}}$\\
Set $\varrho^{(\breve{w} + 1)} = \tilde{\varrho}_{\breve{w}}$ \\
$\flat^{(\breve{w}+1)} = p_{\varrho^{(\breve{w} + 1)}}(\flat^{(\breve{w})}) \coloneqq\text{prox}_{\text{sg} l_1, \tilde{\eta}, \zeta} \big(\flat^{(\breve{w})} - \varrho^{(\breve{w} + 1)} \nabla f(\flat^{(\breve{w})}) \big)$ \\
\If{$\dfrac{\norm{\flat^{(\breve{w}+1)} - \flat^{(\breve{w})}}_2}{\varrho^{(\breve{w} + 1)}} < tol$}
	{
    \textbf{Return:} {$\flat^{\breve{w}}$}
	}
}
\end{algorithm}

>>>>

Now, for the update of $\beta_k^{(w)}$ for $k=0, \ldots, K-1$, we similarly denote
\begin{align*}
  R^{(w)}_{n,k}(\beta_k) = -n^{-1} \sum_{i=1}^n \hat \pi_{ik}^{\theta^{(w)}} \Big[ &\delta_i \hat \E_{\theta^{(w)}}[ \tilde g_{t_i, \gamma_k^{(w)}, \beta_k}^2(b_i) | t_i, \delta_i, y_i] \\
  & - \sum_{j=1}^J \lambda_0^{(w)}(t_j) \hat \E_{\theta^{(w)}}[ \tilde g_{i, t_j, {\gamma_k^0}^{(w)}, \gamma_k^{(w)}, \beta_k}^1(b_i) | t_i, \delta_i, y_i] \ind{\{t_j\leq t_i\}} \\
  & + (y_i \odot \Phi_i^{(w)})^\top \hat \E_{\theta^{(w)}}[ M_{ik} | t_i, \delta_i, y_i] \\
  & - \hat \E_{\theta^{(w)}}[ c_{\phi^{(w)}}(M_{ik}) | t_i, \delta_i, y_i] \Big] 
\end{align*}
based on the quantities involved in $\cQ_n(\theta, \theta^{(w)})$ depending on $\beta_k$, with $\tilde g_{i, t_j, {\gamma_k^0}^{(w)}, \gamma_k^{(w)}, \beta_k}^1$ and $\tilde g_{t_i, \gamma_k^{(w)}, \beta_k}^2$ defined in~\eqref{eq:g-tilde-1-2}. The integration over the survival process has been replaced with a finite sum over the process evaluated at the distinct event times $(t_1, \ldots, t_J)$ with $J \in~\N_+$, since $\lambda_0^{(w)}$ defined in~\eqref{eq:update_lambda_0} is always zero except at observed event times.
The update for $\beta_k^{(w)}$ therefore requires to solve the following convex minimization problem
\begin{equation}
  \label{eq:minimization-beta_k}
  \beta_k^{(w+1)} \in \argmin_{\beta_k \in \R^q} R^{(w)}_{n,k}(\beta_k) + \zeta_{2,k} \norm{\beta_k}_{\text{sg} l_1, \tilde{\eta}}.
\end{equation}
The previous trick we used to handle the non-differentiability of the $\ell_1$ part in the elastic net is no longer sufficient for~\eqref{eq:minimization-beta_k}, since the $\ell_2$ norm in the sparse group lasso is not squared, leading to a non-differentiability at 0 (which is precisely the criterion exploited by the penalty to set some groups of coefficients to exactly 0~\citep{simon2013sparse}). Then, we choose to solve problem~\eqref{eq:minimization-beta_k} using the proximal gradient descent~\citep{boyd2004convex} described in Algorithm~\ref{alg:ISTA}, where we first define
\[p_{\varrho}(y) = \argmin_u \Big\{\dfrac{1}{2\varrho} \left\lVert u - \big(y - \varrho  \nabla f(y) \big)\right\rVert_2^2 + \zeta \norm{u}_{\text{sg} l_1, \tilde{\eta}} \Big\}.\]

@@

\begin{algorithm}[H]
\DontPrintSemicolon
\caption{Proximal gradient descent ISTA$_{\text{sg} l_1}$($v$, $f$, $\tilde \eta$, $\zeta$)}
\label{alg:ISTA}
\KwInput{A vector $v$, a function $f$ (its gradient $\nabla f$), hyper-parameters $\tilde \eta$, $tol$ and $\zeta$}
\KwOutput{$\flat^{\breve{w}^{max}}$}
Initialize $\flat^{(0)}=v$, $\varrho^0 > 0$, some $\epsilon > 1$ \\
\For{$\breve{w} = 0,\ldots,$ until convergence}    
{Backtracking line search : find the smallest nonnegative integer $i^{\breve{w}}$ such that
$f\big(p_{\tilde{\varrho}_{\breve{w}}}(\flat^{(\breve{w})})\big) \leq f(\flat^{(\breve{w})}) + \big<\big(p_{\tilde{\varrho}_{\breve{w}}}(\flat^{(\breve{w})}) - \flat^{(\breve{w})}\big), \nabla f(\flat^{(\breve{w})})\big> + \dfrac{1}{2\tilde{\varrho}_{\breve{w}}}\norm{p_{\tilde{\varrho}_{\breve{w}}}(\flat^{(\breve{w})}) - \flat^{(\breve{w})}}_2^2$ 
with $\tilde{\varrho}_{\breve{w}} = \epsilon^{i^{\breve{w}}}\varrho^{\breve{w}}$\\
Set $\varrho^{(\breve{w} + 1)} = \tilde{\varrho}_{\breve{w}}$ \\
$\flat^{(\breve{w}+1)} = p_{\varrho^{(\breve{w} + 1)}}(\flat^{(\breve{w})}) \coloneqq\text{prox}_{\text{sg} l_1, \tilde{\eta}, \zeta} \big(\flat^{(\breve{w})} - \varrho^{(\breve{w} + 1)} \nabla f(\flat^{(\breve{w})}) \big)$ \\
\If{$\dfrac{\norm{\flat^{(\breve{w}+1)} - \flat^{(\breve{w})}}_2}{\varrho^{(\breve{w} + 1)}} < tol$}
	{
    \textbf{Return:} {$\flat^{\breve{w}}$}
	}
}
\end{algorithm}
